---
output:
  html_document:
    toc: true
    toc_float: true
---

```{r intro, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="hide") 
knitr::opts_knit$set(root.dir = "C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/")

date<-as.Date(Sys.time(	), format='%d%b%Y')
time<-Sys.time()

suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(readxl)))
suppressMessages(library(plotly))
suppressMessages(library(Matrix))
suppressWarnings(suppressMessages(library(zoo)))
suppressWarnings(suppressMessages(library(RColorBrewer))) 

```

```{r dataByDay}
dta<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/KCDC_data.xlsx", sheet = "ByDayNewCases")
str(dta)

    #table(dta$group)
    #table(dta$grouplabel)

dtanewcasebyday<-dta %>%
    filter(is.na(date)==F) %>%
    select(Date_PressRelease, date, newconfirmedcases) %>%
    mutate(date=as.Date(date))

write.csv(dtanewcasebyday, "KCOVID/KCDC_data_exported/dtanewcasebyday.csv")
```

```{r dataByAgeSex}
dta<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/KCDC_data.xlsx", sheet = "ByAgeSex")

    #table(dta$group)
    #table(dta$grouplabel)
dtabyagesex<-dta %>%
    select(Date_PressRelease, group, grouplabel, confirmedcases, deaths) %>%
    mutate(
        date=as.Date(Date_PressRelease, "%m/%d/%y"), 
        deaths=as.numeric(deaths), 
        deaths=ifelse(is.na(deaths)==T, 0, deaths),
        grouplabel=ifelse(grepl("bove", grouplabel) == T, "80-", grouplabel),
        agegroup=grouplabel, 
        agegroup=ifelse(grouplabel=="80-", "80+", agegroup),
        age=sapply(strsplit(grouplabel,"-"), `[`, 1), 
        grouplabel=sapply(strsplit(grouplabel,"-"), `[`, 1), 
        sex=grouplabel
        ) %>%
    arrange(Date_PressRelease, group, grouplabel)

dtapopage<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/WPP2019_KoreaPop.xlsx", sheet="FiveYear")%>%    
    filter(is.na(pop10)==F) %>%
    mutate(
        age=as.character(Age), 
        grouplabel=age, 
        pop=pop10
        ) %>%
    select(grouplabel, pop)

dtapopsex<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/WPP2019_KoreaPop.xlsx", sheet="Sex")%>%
    filter(is.na(pop)==F) %>%
    select(grouplabel, pop) 
    
dtapop<-rbind(dtapopsex, dtapopage)

    #table(dtapop$grouplabel)
    #table(dtabyagesex$grouplabel)

dtabyagesex<-left_join(dtabyagesex, dtapop, by = "grouplabel") %>%
    mutate(
        incidence=round(100*confirmedcases/pop, 1), #per 100,000 pop
        cfr=round(100*deaths/confirmedcases, 2),
        latest=date==max(date), 
        initial=date==min(date) 
        ) 

dtabyage<-dtabyagesex %>%filter(group=="age")%>% 
    arrange(Date_PressRelease, age)

dtabysex<-dtabyagesex%>%filter(group=="sex")

#export to CSV for the paper 
write.csv(dtabyagesex, "KCOVID/KCDC_data_exported/dtabyagesex.csv")
write.csv(dtabyage, "KCOVID/KCDC_data_exported/dtabyage.csv")
```

```{r dataSummaryByDay}
#data for all trend data: testing, incidnece, mortality 
dtapop<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/WPP2019_KoreaPop.xlsx", sheet="FiveYear") %>%
    filter(is.na(pop10)==F)  
totalpop<-round(sum(dtapop$pop10)) #total population in thousand

dta<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/KCDC_data.xlsx", sheet = "SummaryByDay")
    
names(dta)<- tolower(names(dta))   
str(dta)
dtabyday<-dta %>%
    filter(is.na(total)==F) %>%
    mutate(
        totalpop<-totalpop, 
        date=as.Date(date), 
        confirmed_discharged= ifelse(is.na(confirmed_discharged)==T, 0, confirmed_discharged),
        confirmed_deceased= ifelse(is.na(confirmed_deceased)==T, 0, confirmed_deceased),
        confirmed_isolated= ifelse(is.na(confirmed_isolated)==T, confirmed_total, confirmed_isolated),
        confirmed_total= ifelse(is.na(confirmed_total)==T, confirmed_discharged + confirmed_isolated + confirmed_deceased , confirmed_total),
        piu_total= ifelse(is.na(piu_total)==T, piu_beingtested+piu_testednegative, piu_total)
        ) %>%
    arrange(date) %>%
    mutate(
        test=lead(total)-total, #new tests in each day
        testpositive=lead(confirmed_total)-confirmed_total, #new positive in each day
        cfr=round(100*confirmed_deceased / confirmed_total, 2), #cumulative CFR (%)
        incidence=round(100*confirmed_total/totalpop, 1) #cumulative incidence per 100,000 population 
        )

#export to CSV for the paper 
write.csv(dtabyday, "KCOVID/KCDC_data_exported/dtabyday.csv")
```

#####__COVID-19 CFR in South Korea: further investigation__ 

(Updated: `r time` EDT)  

---

####__1. Trends of testing and new cases__

```{r plottest, results="asis", fig.align="left", out.width="800px"}
dta<-left_join(dtabyday, dtanewcasebyday, by = "date")%>%
    arrange(date)%>%
    mutate(positiverate=round(100*(newconfirmedcases/test), 1),
            positiveratesmooth =round(c(NA,NA,NA,rollmean(positiverate, 7),NA,NA,NA), 5), 
            positiveratesmooth =ifelse(positiveratesmooth<0,  0, positiveratesmooth)
            )

plot_ly(dta, x=~date, 
        y=~test, name="Tests per day",
        type='bar',  
        marker=list(color = c( "#8DB9D9"), opacity=1, size=2)
        ) %>%
    add_lines(
        y=~positiveratesmooth, name="Positive rate", 
        marker=list(color = "black"),
        line=list(color = "black"),
        yaxis='y2')%>%
    add_segments(x = min(dta$date), xend = max(dta$date), 
                 y = 5, yend = 5,yaxis='y2', 
                 marker = list(color = "gray",size = 2, dash="dot"),
                 line= list(color = "gray"),
                 showlegend=FALSE)%>%
    layout(
        title = c("Trend of daily tests and new cases"),
        xaxis = list(title = "", tickfont = list(size=10), showgrid = FALSE), 
        yaxis = list(title = "Number of tests", 
                     range=c(0, 20000),
                     side="left", showgrid = FALSE ),
        yaxis2 = list(title = "Positive rate (%)", 
                      range=c(0, 20),
                      overlaying='y',  
                      side="right", showgrid = FALSE),
        margin = list(b = 100, r=100), 
        legend=list(orientation="v", xanchor = "left", yanchor = "center", 
                    x = 0.05, y = 0.9) 
        
        )    
```

---

```{r plotnewtrend, results="asis", fig.align="left", out.width="800px"}
latestsummary<- dtanewcasebyday %>% 
    select(date) %>% summarize_all(funs(max))

dtafig<-dtanewcasebyday%>%
    mutate(
        newconfirmedcasessmooth =c(NA,NA,NA,rollmean(newconfirmedcases, 7),NA,NA,NA), 
        newconfirmedcasessmooth = round(newconfirmedcasessmooth, 1),
        startdate=as.Date("2020-02-23"),
        peakdate=as.Date("2020-03-03"), 
        enddate=as.Date("2020-03-12") 
        )
              
shapelist<-list(
            list(
                type = "rect",  fillcolor = "#f03b20", 
                line = list(color = "#f03b20"), opacity = 0.2,
                x0=min(dtafig$startdate), x1=max(dtafig$peakdate, na.rm = TRUE),xref = "x",
                y0=0,y1=1000, yref = "y") , 
            list(
                type = "rect",  fillcolor = "#fd8d3c", 
                line = list(color = "#fd8d3c"), opacity = 0.2,
                x0=min(dtafig$peakdate, na.rm = TRUE), x1=max(dtafig$enddate, na.rm = TRUE),xref = "x",
                y0=0,y1=1000, yref = "y" ),
            list(
                type = "rect",  fillcolor = "#fed976", 
                line = list(color = "#fed976"), opacity = 0.2,
                x0=max(dtafig$enddate, na.rm = TRUE), x1=max(dtafig$date),xref = "x",
                y0=0,y1=1000, yref = "y" )
            )
  
plot_ly(dtafig, x=~date, 
        y=~newconfirmedcases, name="New cases", 
        type='bar', 
        marker=list(color = c( "lightgray"), size = 1),
        line=list(color = c( "lightgray"))
        ) %>%
    add_trace(
        y=~newconfirmedcasessmooth, name="New cases, 7-day rolling average",
        type='scatter',mode = 'lines', 
        marker=list(color = c( "black"), size = 1),
        line=list(color = c( "black"))
        )%>%
    add_segments(x = as.Date("2020-01-20"), xend = max(date), 
                 y = 154, yend = 154, 
                 marker = list(color = "black", size = 2.2),
                 line= list(color = "black", dash = 'dot'),
                 showlegend=FALSE)%>%
    add_annotations(x = as.Date("2020-04-05"), y = 250, yaxis='y',
                    text="Number of new cases",
                    showarrow = FALSE, 
                    font = list(color = "black", size = 12))%>%        
    add_annotations(x = as.Date("2020-04-05"), y = 220, yaxis='y',
                    text="at the start of the first wave",
                    showarrow = FALSE, 
                    font = list(color = "black", size = 12))%>%        
    add_annotations(x = as.Date("2020-04-05"), y = 190, yaxis='y',
                    text="(154 per day)",
                    showarrow = FALSE, 
                    font = list(color = "black", size = 12))%>%        
    layout(
        shapes=shapelist, 
        title = c("Trend of new cases before, during, and after the first wave"),
        xaxis = list(title = "", tickfont = list(size=10), showgrid = FALSE), 
        yaxis = list(title = "New cases", 
                     range=c(0,1000)),
        margin = list(b = 100, r=100), 
        legend=list(orientation="h", xanchor = "center", yanchor = "center", 
                    x = 0.5, y = -0.1) 
        )

```

_Note: Three phases (red/orange/yellow shades) based on the trend of new concirmed cases. [See here for more information on that classification](https://medium.com/@yj.choi.isquared/flattening-covid-19-curve-in-south-korea-and-comparison-among-oecd-countries-singapore-and-taiwan-ae211a5645c9)_.

```{r plotincidencecfrtrend, results="asis", fig.align="left", out.width="800px"}
dtafig<-dtabyday%>%select(date, cfr, incidence)
#dta<-dtabyday%>%filter(cfr>0)%>%select(date, cfr, incidence)
dtafig$incidence<-dtafig$incidence/10

plot_ly(dtafig, x=~date, y=~incidence, 
        type = 'bar', name="Incidence rate (per million population)",
        marker = list(color = "rgb(204,204,204)")) %>%
        add_lines(y = ~cfr, name = "Case fatality rate (per 100 cases)" ,
                  marker = list(color = "rgb(253,141,60)")) %>%
        layout(
            title = "Trends of cumulative incidence and case fatality rates by day",
            yaxis = list(title="", range=c(0, 2.5)),  
            xaxis = list(  
                         tickfont = list(size=10)), 
            legend = list(orientation = 'v',x = 0.05, y = 0.9) 
            )
```

---

####__2. Decomposing age composition and mortality risk by age__ 

#####__2.1. Trends of age-specific incidence rates and CFR__

```{r plotincidencetrendbyage, results="asis", fig.align="left", out.width="700px"}

dtasorted<-dtabyage %>%
    select(date, age, incidence) %>%
    mutate(
        age=paste0("incidence",age)
    ) %>%
    arrange(date, age)
#head(dtasorted, 15)

dta<-dtasorted %>% 
    spread(age, incidence, fill = NA, convert = FALSE) 
#head(dta, 20)
#str(dta)

fig <- plot_ly(dta, x=~date, y=~incidence0, name = "0-9", 
               type = 'scatter', mode='lines',
               line = list(color = c( "rgb(204,236,230)"))) %>% 
    add_trace(y = ~incidence10, name = "10-19", line = list(color = c( "rgb(153,216,201)"))) %>% 
    add_trace(y = ~incidence20, name = "20-29", line = list(color = c( "rgb(102,194,164)"))) %>% 
    add_trace(y = ~incidence30, name = "30-39", line = list(color = c( "rgb(65,174,118)"))) %>% 
    add_trace(y = ~incidence40, name = "40-49", line = list(color = c( "rgb(35,139,69)"))) %>% 
    add_trace(y = ~incidence50, name = "50-59", line = list(color = c( "rgb(0,109,44)"))) %>% 
    add_trace(y = ~incidence60, name = "60-69", line = list(color = c( "rgb(0,68,27)"))) %>% 
    add_trace(y = ~incidence70, name = "70-79", line = list(color = c("#FF9900"), width=4)) %>% 
    add_trace(y = ~incidence80, name = "80+", line = list(color = c("#CC0000"), width=4)) %>% 
    layout(
        title ="Trends of cumulative incidence rate by age group",
        yaxis = list(title = "Number of confirmed cases (per 100000 population)",
                     titlefont=list(size=12),
                     type="log"), 
        legend = list(font=list(size=10)) 
        ) 

fig
```

```{r plotmortalitytrendbyage, results="asis", fig.align="left", out.width="800px"}

panel <- . %>% 
    plot_ly(x=~date, 
            y=~cfr, name="CFR",
            type = 'scatter', mode = 'lines',
            line=list(color = c( "red"))
            ) %>%
    add_annotations(
        text = ~unique(agegroup),
        x = 0.5, y = 0.9, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%
    layout(
        showlegend = FALSE,
        title=c("Trend of case fatality rate by age group"),
        xaxis=list(title = "Date", tickfont = list(size=8)),
        yaxis=list(title = "Case fatality rate (%)", 
                   range=c(0.2, upperrange)
                   )
        
        )

upperrange<-round(max(dtabyage$cfr)+1) 

dtabyage%>%filter(age>=40)%>%
    group_by(agegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 1, shareY = TRUE) 
```

```{r plottrendbyage, results="asis", fig.align="left", out.width="800px"}

panel <- . %>% 
    plot_ly(x=~date, 
            y=~incidence, name="Incidence per 100,000",
            type = 'scatter', mode = 'lines',
            line=list(color = c( "blue"))
            ) %>%
    add_lines(
            y=~cfr, name="CFR",
            line=list(color = "red"),
            yaxis='y2'
            ) %>%
    add_annotations(
        text = ~unique(agegroup),
        x = 0.5, y = 0.95, xref = "paper", yref = "paper",    
        xanchor = "center", yanchor = "bottom", showarrow = FALSE,
        font = list(size = 12)
        ) %>%
    layout(
        showlegend = FALSE,
        title=c("Trend of incidence and case fatality rates by age group"),
        xaxis=list(title = "Date", tickfont = list(size=8)),
        yaxis = list(title = "Incidence rate (per 100,000 population)", 
                     range=c(0, 30),
                     side="left"),
        yaxis2=list(title = "Case fatality rate (%)", 
                    range=c(0, 30),
                    side="right")
                   )

dtabyage%>%filter(age>=40)%>%
    group_by(agegroup) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 1, shareY = TRUE) 
```



#####__2.2. Composition by age: population, confirmed COVID-19 cases, and COVID-19 deaths__

```{r plotagecomposition, results="asis", fig.align="left", out.width="800px"}
#Population
temp1 <- dtabyage%>%select(date, age, agegroup, pop)%>%
    filter(date==max(date)) %>%     
    mutate(
        total=sum(pop), 
        pct=round(100*pop/total,1)  ,
        group="1. Population"  
    )%>%select(age, agegroup, group, pct)

#cases
temp <- dtabyage%>%select(date, confirmedcases)%>%
    group_by(date)%>%summarize_all(funs(sum))%>%
    mutate(total=confirmedcases)%>%select(date, total)

temp2<-left_join(dtabyage, temp, by = "date") %>% 
    select(date, age, agegroup, confirmedcases, total) %>% 
    filter(date==max(date)) %>%     
    mutate(
        pct=round(100*confirmedcases/total,1)    ,
        group="2. Cases"
    )%>%select(age, agegroup, group, pct)


#deaths
temp <- dtabyage%>%select(date, deaths)%>%
    group_by(date) %>%summarize_all(funs(sum))%>%
    mutate(total=deaths)%>%select(date, total)

temp3<-left_join(dtabyage, temp, by = "date") %>% 
    select(date, age, agegroup, deaths, total) %>% 
    filter(date==max(date)) %>%     
    mutate(
        pct=round(100*deaths/total,1)   ,
        group="3. Deaths"
    )%>%select(age, agegroup, group, pct)

# merge all three
dtafig<-rbind(temp1, temp2, temp3) 

dtafig %>% 
    plot_ly( x=~group, y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>% 
    layout(
        title ="Age pattern of population, cases, and deaths",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 



```


#####__2.3. Changes in composition by age: confirmed COVID-19 cases and COVID-19 deaths__

```{r plotagecompositiontrend, results="asis", fig.align="left", out.width="800px"}
#cases
temp <- dtabyage%>%select(date, confirmedcases)%>%
    group_by(date)%>%summarize_all(funs(sum))%>%
    mutate(total=confirmedcases)%>%select(date, total)

dtafig<-left_join(dtabyage, temp, by = "date") %>% 
    select(date, agegroup, confirmedcases, total) %>% 
    mutate(
        pct=round(100*confirmedcases/total,1)    
    )

figcases<-dtafig %>% group_by(agegroup) %>% 
    arrange(date) %>%
    plot_ly( x = ~date, y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>%
    layout(
        title ="Age pattern of confirmed cases over time",
        yaxis = list(title = "Percent of confirmed cases",
                     titlefont=list(size=12)), 
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

#deaths
temp <- dtabyage%>%select(date, deaths)%>%
    group_by(date)%>%summarize_all(funs(sum))%>%
    mutate(total=deaths)%>%select(date, total)

dtafig<-left_join(dtabyage, temp, by = "date") %>% 
    select(date, agegroup, deaths, total) %>% 
    mutate(
        pct=round(100*deaths/total,1)    
    )

figdeaths<-dtafig %>% group_by(agegroup) %>% 
    arrange(date) %>%
    plot_ly( x = ~date, y = ~pct, type = "bar",
             color= ~agegroup, 
             colors = brewer.pal(length(unique(dtafig$agegroup)),
                                "Spectral")
             ) %>%
    layout(
        title ="Age pattern of deaths over time",
        yaxis = list(title = "Percent of deaths",
                     titlefont=list(size=12)), 
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 


subplot(style(figcases, 
              showlegend = FALSE
              ), 
        style(figdeaths, 
              legend = list(x = 1.05, y = 0.5, font=list(size=10))
              ),
        nrows=1, margin = 0.05, shareY = FALSE, titleY = TRUE
        )



```

---

####__3. Decomposing sex composition and mortality risk by sex__

#####__3.1. Trends of sex-specific incidence rates and CFR__ (coming soon)

#####__3.2. Composition by sex: population, confirmed COVID-19 cases, and COVID-19 deaths__

```{r plotsexcomposition, results="asis", fig.align="left", out.width="800px"}
#Population
temp1 <- dtabysex%>%select(date, sex, pop)%>%
    filter(date==max(date)) %>%     
    mutate(
        total=sum(pop), 
        pct=round(100*pop/total,1)  ,
        group="1. Population"  
    )%>%select(sex, group, pct)

#cases
temp <- dtabysex%>%select(date, confirmedcases)%>%
    group_by(date)%>%summarize_all(funs(sum))%>%
    mutate(total=confirmedcases)%>%select(date, total)

temp2<-left_join(dtabysex, temp, by = "date") %>% 
    select(date, sex, confirmedcases, total) %>% 
    filter(date==max(date)) %>%     
    mutate(
        pct=round(100*confirmedcases/total,1)    ,
        group="2. Cases"
    )%>%select(sex, group, pct)


#deaths
temp <- dtabysex%>%select(date, deaths)%>%
    group_by(date) %>%summarize_all(funs(sum))%>%
    mutate(total=deaths)%>%select(date, total)

temp3<-left_join(dtabysex, temp, by = "date") %>% 
    select(date, sex, deaths, total) %>% 
    filter(date==max(date)) %>%     
    mutate(
        pct=round(100*deaths/total,1)   ,
        group="3. Deaths"
    )%>%select(sex, group, pct)

# merge all three
dtafig<-rbind(temp1, temp2, temp3) 

dtafig %>% 
    plot_ly( x=~group, y = ~pct, type = "bar",
             color= ~sex, 
             colors = brewer.pal(length(unique(dtafig$sex)),
                                "Spectral")
             ) %>% 
    layout(
        title ="Age pattern of population, cases, and deaths",
        yaxis = list(title = "Percent",
                     titlefont=list(size=12)), 
        yaxis = list(title = ""),
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 



```

#####__3.3. Changes in composition by sex: confirmed COVID-19 cases and deaths__

```{r plotsexcompositiontrend, results="asis", fig.align="left", out.width="800px"}
### cases
temp <- dtabysex%>%filter(group=="sex")%>%
    select(date, confirmedcases)%>%
    group_by(date)%>%summarize_all(funs(sum))%>%
    mutate(total=confirmedcases)%>%select(date, total)

dtafig<-left_join(dtabysex, temp, by = "date") %>% 
    select(date, sex, confirmedcases, total) %>% 
    mutate(
        pct=round(100*confirmedcases/total,1)    
    )

figcases<-dtafig %>% group_by(sex) %>% 
    arrange(date) %>%
    plot_ly( x = ~date, y = ~pct, type = "bar",
             color= ~sex, 
             colors = brewer.pal(length(unique(dtafig$sex)),
                                "Spectral")
             ) %>%
    layout(
        title ="Sex pattern of confirmed cases over time",
        yaxis = list(title = "Percent of confirmed cases",
                     titlefont=list(size=12)), 
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

### deaths
temp <- dtabysex%>%filter(group=="sex")%>%
    select(date, deaths)%>%
    group_by(date)%>%summarize_all(funs(sum))%>%
    mutate(total=deaths)%>%select(date, total)

dtafig<-left_join(dtabysex, temp, by = "date") %>% 
    select(date, sex, deaths, total) %>% 
    mutate(
        pct=round(100*deaths/total,1)    
    )

figdeaths<-dtafig %>% group_by(sex) %>% 
    arrange(date) %>%
    plot_ly( x = ~date, y = ~pct, type = "bar",
             color= ~sex, 
             colors = brewer.pal(length(unique(dtafig$sex)),
                                "Spectral")
             ) %>%
    layout(
        title ="Sex pattern of deaths over time",
        yaxis = list(title = "Percent of deaths",
                     titlefont=list(size=12)), 
        legend = list(font=list(size=10)),
        barmode = 'stack'
        ) 

subplot(style(figcases, 
              showlegend = FALSE
              ), 
        style(figdeaths, 
              legend = list(x = 1.05, y = 0.5, font=list(size=10))
              ),
        nrows=1, margin = 0.05, shareY = FALSE, titleY = TRUE
        )

```

---

__Data source__: KCDC's [daily press release](https://www.cdc.go.kr/board/board.es?mid=&bid=0030), and UN [World Population Prospects 2019 Revision](https://population.un.org/wpp/)   
_Note_: Age-specific mortality data were first released on March, 3rd, and it has been updated daily since March 11th. 

---

<p style="color:gray">
More questions and data coming soon, as I continue digging the awesome KCDC website...    
See [GitHub](https://github.com/yoonjoung/COVID19_SouthKorea) for data, code, and more information.  
For typos, errors, and questions, contact me at [www.isquared.global](https://www.iSquared.global/YJ). 

_Making Data Delicious, One Byte at a Time_, in good times and bad times.</p>
