---
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---
```{r intro, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="hide") 
knitr::opts_knit$set(root.dir = "C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/")

date<-as.Date(Sys.time(	), format='%d%b%Y')
time<-Sys.time()

suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(readxl)))
suppressMessages(library(plotly))
suppressMessages(library(Matrix))

```

```{r dataByRegion}
dtapop<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/WPP2019_KoreaPop.xlsx", sheet="FiveYear") %>%
    filter(is.na(pop10)==F)  
totalpop<-round(sum(dtapop$pop10)) #total population in thousand

dtapopregion<-read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/Citypopulation_SouthKorea.xlsx") 

dtapopregion$pop=dtapopregion[,10]
#table(dtapopregion$region)

dtapopregion<-dtapopregion %>%
    select(region, pop) %>%
    mutate(
        region=ifelse(region=="Seoul", "gg", region), 
        region=ifelse(region=="Incheon", "gg", region),  
        region=ifelse(region=="Gyeonggi-do", "gg", region),
        region=ifelse(region=="Gangwon-do", "gw", region),
        region=ifelse(region=="Chungcheongbuk-do", "cb", region),        
        region=ifelse(region=="Chungcheongnam-do", "cn", region),
        region=ifelse(region=="Daejeon", "cn", region),
        region=ifelse(region=="Sejong", "cn", region),
        region=ifelse(region=="Daegu", "gb", region),
        region=ifelse(region=="Ulsan", "gb", region),
        region=ifelse(region=="Gyeongsangbuk-do", "gb", region),
        region=ifelse(region=="Busan", "gn", region),
        region=ifelse(region=="Gyeongsangnam-do", "gn", region),
        region=ifelse(region=="Jeollabuk-do", "jb", region),
        region=ifelse(region=="Gwangju", "jn", region),
        region=ifelse(region=="Jeollanam-do", "jn", region),
        region=ifelse(region=="Jeju-do", "jj", region)
    ) %>%
    group_by(region) %>% summarize_all(funs(sum)) %>%
    mutate(
        total=sum(pop),
        pct=pop/total,#pop in 1000
        pop=totalpop*pct,#total pop from WPP 2019 revision
        total2=sum(pop)
        ) %>%
    select(region, pop) %>%arrange(region)
#head(dtapopregion, 10)

dta<-as.data.frame(read_excel("C:/Users/YoonJoung Choi/Dropbox/0 Project/COVID19_SouthKorea/KCDC_data.xlsx", sheet = "ByRegion"))

names(dta) <- gsub(" ", "", names(dta))
names(dta)<- tolower(names(dta))   

colnames(dta[,8:25])

###make sure the region/city columns are all numeric 
###replace missing with 0 in region/city columns 
for (i in 8:25){
    dta[,i]<-as.numeric(dta[,i])    
    for (j in 1:nrow(dta)){
        if (is.na(dta[j,i])==TRUE){
            dta[j,i]<-0    
        }
    }
}

dtabyregiondetail<-dta %>%
    mutate_each(funs(tolower), status) %>%
    filter(metric=="cumulative cases") %>%
    filter(status!="" & status!="shincheonji") %>%
    mutate(
        status=ifelse(status=="sub-total", "total", status),    
        status=ifelse(status=="subtotal", "total", status) 
    ) %>%
    arrange(date, status, time) %>%
    mutate(
        gg=seoul+ incheon+ gyeonggi, 
        gw=gangwon,
        cb=chungbuk,
        cn=daejeon+ sejong+ chungnam,   
        gb=daegu+ ulsan+ gyeongbuk,            
        gn=busan+ gyeongnam,            
        jb=jeonbuk,          
        jn=gwangju+ jeonnam,           
        jj=jeju,             
        airport=airportscreening, 
        TOTAL=gg+gw+cn+cb+gb+gn+jb+jn+jj+airport
    ) %>%
    select(date, time, status, gg,gw,cn,cb,gb,gn,jb,jn,jj,airport, 
           total, TOTAL)%>% 
    group_by(date)%>%
    mutate(
        tworeports =sd(time)!=0, 
        early =min(time)
        )%>%ungroup()%>%
    filter(tworeports==0 | (tworeports==1 & early==time))%>%
    select(-time, -tworeports, -early, -total)%>%
    arrange(date)

### TOTAL cumulative cases
dtabyregiontotal<-dtabyregiondetail %>% 
    filter(status=="total") %>% select(-status) %>% arrange(date)

dtabyregion<-dtabyregiontotal %>% 
    gather(region, cases, gg:airport, convert = FALSE) %>% 
    arrange(region, date) %>% 
    mutate(
        newcases=cases-lag(cases),
        newcases=ifelse(region!=lag(region), NA, newcases )
    )
    
dtabyregion<-left_join(dtabyregion, dtapopregion, by = "region") 

dtabyregion<-dtabyregion %>% 
    mutate(
        incidence =round(100000* cases/(pop*1000), 1), #per 100000 pop
        pctcases = round(100*cases/TOTAL, 1)#% region cases out of total
    )

### Deaths 
dtabyregiondeaths<-dtabyregiondetail %>% 
    filter(status=="deceased") %>% select(-status ) %>% arrange(date)%>%
    mutate(TOTALdeaths=TOTAL)%>% select(-TOTAL )

dtabyregiondeaths<-dtabyregiondeaths %>% 
    gather(region, deaths, gg:airport, convert = FALSE) %>% 
    arrange(region, date) 

dtabyregion<-left_join(dtabyregion, dtabyregiondeaths, by = c("region", "date")) 

dtabyregion<-dtabyregion %>% 
    mutate(
        cfr =round(100* deaths/cases, 1), #CFR %
        region=ifelse(region=="gg", "Gyeonggi", region),
        region=ifelse(region=="gw", "Gangwon", region),
        region=ifelse(region=="cb", "Chungbuk", region),
        region=ifelse(region=="cn", "Chungnam", region),
        region=ifelse(region=="jb", "Jeonbuk", region),
        region=ifelse(region=="jn", "Jeonnam", region),
        region=ifelse(region=="gb", "Gyungbuk", region),
        region=ifelse(region=="gn", "Gyungnam", region),
        region=ifelse(region=="jj", "Jeju", region),
        region=ifelse(region=="airport", "Airport", region)
    )%>%
    arrange(date, region)
```

#####__COVID-19 regional variation in South Korea__ 

(Updated: `r time` EDT)  

* This shows supplementary figures for [COVID-19 in South Korea - and potential implications for us (wherever we are)](https://rpubs.com/YJ_Choi/COVID19_SouthKorea)      
* All data are from press release from [KCDC](https://www.cdc.go.kr/) - available to the public in both Korean and English. KCDC has released regional data since March 7th.    
* Special cities are included in provinces where they are located: Daejeon and Sejong in Chungnam province; Seoul and Inchon in Gyeonggi province; Daegu and Ulsan are included in Gyungbuk province; Busan in Gyungnam province; and Gwangju in Jeonnam province.   

---

#####__1. Cumulative cases by region__ 
```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%filter(date==max(date))

dtafig$region<-factor(dtafig$region, 
                    levels = unique(dtafig$region) 
                    [order(dtafig$cases, decreasing = TRUE)])

fig1<-plot_ly(dtafig, x=~region, y=~cases, type = 'bar',
              marker = list(color = c('rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(49,130,189)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)')))%>% 
    layout(
            title = c("COVID-19 cases by region"),
            xaxis = list(title = "Date",  
                         autotick = FALSE,
                         showticklabels = TRUE, 
                         tickfont = list(size=10), 
                         tickangle=-90),
            yaxis = list(title = "Number of confirmed cases"), 
            legend = list(x = 1.2, y = 0.9, 
                          title=list(text='Provinces*'))
            )

dtafig<-dtabyregion%>%filter(date==max(date))%>%filter(region!="Gyungbuk")

dtafig$region<-factor(dtafig$region, 
                    levels = unique(dtafig$region) 
                    [order(dtafig$cases, decreasing = TRUE)])

fig2<-plot_ly(dtafig, x=~region, y=~cases, type = 'bar',
              marker = list(color = c('rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)')))%>% 
    layout(
            title = c("COVID-19 cases by region - except GB"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Number of Confirmed cases"), 
            legend = list(x = 1.2, y = 0.9, 
                          title=list(text='Provinces*'))
            )

subplot(fig1, fig2, nrows=1) %>%
        layout(  title ="Cumulative cases by region: w/ vs. wo/ GB" , 
                 showlegend = FALSE )

```

---

#####__2. Cumulative incidence by region__ 
```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%filter(date==max(date))

dtafig$region<-factor(dtafig$region, 
                    levels = unique(dtafig$region) 
                    [order(dtafig$incidence, decreasing = TRUE)])

fig1<-plot_ly(dtafig, x=~region, y=~incidence, type = 'bar',
              marker = list(color = c('rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(49,130,189)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)')))%>% 
    layout(
            title = c("COVID-19 incidence by region"),
            xaxis = list(title = "Date",  
                         autotick = FALSE,
                         showticklabels = TRUE, 
                         tickfont = list(size=10), 
                         tickangle=-90),
            yaxis = list(title = "Confirmed cases per 100,000 population"), 
            legend = list(x = 1.2, y = 0.9, 
                          title=list(text='Provinces*'))
            )

dtafig<-dtabyregion%>%filter(date==max(date))%>%filter(region!="Gyungbuk")

dtafig$region<-factor(dtafig$region, 
                    levels = unique(dtafig$region) 
                    [order(dtafig$incidence, decreasing = TRUE)])

fig2<-plot_ly(dtafig, x=~region, y=~incidence, type = 'bar',
              marker = list(color = c('rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)', 'rgb(107,174,214)')))%>% 
    layout(
            title = c("COVID-19 incidence by region"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Confirmed cases per 100,000 population"), 
            legend = list(x = 1.2, y = 0.9, 
                          title=list(text='Provinces*'))
            )

subplot(fig1, fig2, nrows=1) %>%
        layout(  title ="Incidence rate (per 100,000 population) by region: w/ vs. wo/ GB" , 
                 showlegend = FALSE )

```

---

#####__2.1 Trend of cumulateive incidence rates by region over time__ 
```{r,  fig.align="left", out.width="800px"}
listx<-list(title = "Date",  
                         tickfont = list(size=10))
listy<-list(range=c(0, 7))

dtafig<-dtabyregion%>%filter(region=="Gyeonggi")
figgg<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Gyeonggi")%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Gangwon")
figgw<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Gangwon")%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Chungbuk")
figcb<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Chungbuk" )%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Chungnam")
figcn<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Chungnam")%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Jeonbuk")
figjb<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Jeonbuk" )%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Jeonnam")
figjn<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Jeonnam")%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Gyungbuk")
figgb<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Gyungbuk")%>%
        layout(xaxis=listx, yaxis=list(range=c(0, 150)))

dtafig<-dtabyregion%>%filter(region=="Gyungnam")
figgn<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Gyungnam" )%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Jeju")
figjj<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Jeju" )%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Airport")
figairport<-plot_ly(dtafig, x=~date, y=~incidence, type = 'bar', name="Airport")%>%
        layout(xaxis=listx, yaxis=listy)

subplot(figgb, figgg, figgw, figcn, figcb, figjb, figjn, figgn, figjj, 
        nrows=3, shareX=TRUE, shareY=FALSE) %>%
        layout(
            title ="Trends of cumulative incidence rate (per 100,000) by region",
            showlegend=FALSE, 
            margin = list(b = 100)
            )
```

```{r,  results='asis', fig.align="left", out.width="800px"}

panel <- . %>% 
  plot_ly(x=~date, y=~incidence, type = 'bar', marker=list(color=("'#1F77B4'")) ) %>%
  add_annotations(
    text = ~unique(region),
    x = 0.5, y = 0.9, xref = "paper", yref = "paper",    
    xanchor = "center", yanchor = "bottom", showarrow = FALSE,
    font = list(size = 12)
  ) %>%
  layout(
    showlegend = FALSE,
    xaxis=list(title = "Date",  
             font=list(size=10),
             tickfont = list(size=8) 
             ),
    yaxis=list(title = "incidence rate (per 100,000)"))


dtafig<-dtabyregion%>%filter(region!="Airport")

dtafig%>%
    group_by(region) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 3, shareX = TRUE, shareY = FALSE)%>%
        layout(
            title ="Trends of cumulative incidence rate (per 100,000) by region",
            showlegend=FALSE, 
            margin = list(b = 100)
            )   
```

---

#####__3. Trend of regional composition among confirmed cases over time__
```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion
plot_ly(dtafig, x=~date, y=~pctcases, type = 'bar', color= ~region )%>% 
    layout(
            title = c("Trend of case composition by region"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Percent of total cases"),
            barmode="stack",  
            legend = list(x = 1.05, y = 0.9, 
                          title=list(text='Provinces*'))
            )
```

---

#####__4. Trend of new cases by region over time__
```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%
    mutate(newcases = ifelse(newcases<0, 0, newcases))%>%
    plot_ly(dtafig, x=~date, y=~newcases, 
        type = 'bar', color= ~region ) %>%
        layout(
            title = c("Trend of daily new confirmed COVID-19 cases by region"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Number of new confirmed cases"),
            barmode="stack",  
            legend = list(x = 1.05, y = 0.9, 
                          title=list(text='Provinces*'))
            )
```

```{r,  fig.align="left", out.width="800px"}
listx<-list(title = "Date", tickfont = list(size=8))

listy<-list(range=c(0, 30))
listy2<-list(range=c(0, 15))

dtafig<-dtabyregion%>%filter(region=="Gyeonggi")
figgg<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Gyeonggi")%>%
        layout(xaxis=listx)

dtafig<-dtabyregion%>%filter(region=="Gangwon")
figgw<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Gangwon")%>%
        layout(xaxis=listx, yaxis=listy2)

dtafig<-dtabyregion%>%filter(region=="Chungbuk")
figcb<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Chungbuk" )%>%
        layout(xaxis=listx, yaxis=listy2)

dtafig<-dtabyregion%>%filter(region=="Chungnam")
figcn<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Chungnam")%>%
        layout(xaxis=listx, yaxis=listy)

dtafig<-dtabyregion%>%filter(region=="Jeonbuk")
figjb<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Jeonbuk" )%>%
        layout(xaxis=listx, yaxis=listy2)

dtafig<-dtabyregion%>%filter(region=="Jeonnam")
figjn<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Jeonnam")%>%
        layout(xaxis=listx, yaxis=listy2)

dtafig<-dtabyregion%>%filter(region=="Gyungbuk")
figgb<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Gyungbuk")%>%
        layout(xaxis=listx)

dtafig<-dtabyregion%>%filter(region=="Gyungnam")
figgn<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Gyungnam" )%>%
        layout(xaxis=listx, yaxis=listy2)

dtafig<-dtabyregion%>%filter(region=="Jeju")
figjj<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Jeju" )%>%
        layout(xaxis=listx, yaxis=listy2)

dtafig<-dtabyregion%>%filter(region=="Airport")
figairport<-plot_ly(dtafig, x=~date, y=~newcases, type = 'bar', name="Airport")%>%
        layout(xaxis=listx, yaxis=listy)

subplot(figgb, figgg, figairport, figcn, figgw, figcb, figjb, figjn, figgn, figjj, 
        nrows=3, shareX=TRUE, shareY=FALSE) %>%
        layout(
            title ="Trends of new cases by region",
            legend = list(orientation = 'h', x = 0.2) ,
            margin = list(b = 100)
            )

```

```{r,  results='asis', fig.align="left", out.width="800px"}

panel <- . %>% 
  plot_ly(x=~date, y=~newcases, type = 'bar', marker=list(color=("'#1F77B4'")) ) %>%
  add_annotations(
    text = ~unique(region),
    x = 0.5, y = 0.9, xref = "paper", yref = "paper",    
    xanchor = "center", yanchor = "bottom", showarrow = FALSE,
    font = list(size = 12)
  ) %>%
  layout(
    showlegend = FALSE,
    xaxis=list(title = "Date",  
             font=list(size=10),
             tickfont = list(size=8) 
             ),
    yaxis=list(title = "New cases"))


dtafig<-dtabyregion%>%filter(region!="Airport")%>%
    mutate(newcases = ifelse(newcases<0, 0, newcases))

dtafig%>%
    group_by(region) %>%
    do(p = panel(.)) %>%
    subplot(nrows = 3, shareX = TRUE, shareY = FALSE)%>%
        layout(
            title ="Trends of new cases by region",
            showlegend=FALSE, 
            margin = list(b = 100)
            )   
```

---

#####__5. Number of deaths by region__
```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%filter(cfr>0)%>%filter(region!="Airport")
ncfr<-length(unique(dtafig$region))
```

As of `r date`, COVID-19 deaths have been reported in `r ncfr` provinces. 

```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%filter(date==max(date))%>%filter(region!="Airport")

dtafig$region<-factor(dtafig$region, 
                    levels = unique(dtafig$region) 
                    [order(dtafig$deaths, decreasing = TRUE)])

plot_ly(dtafig, x=~region, y=~deaths, type = 'bar')%>% 
    layout(
            title = c("COVID-19 deaths by region"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Number of deaths"), 
            legend = list(x = 1.2, y = 0.9, 
                          title=list(text='Provinces*'))
            )
```

---

#####__6. Case fatality rate by region__
```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%filter(date==max(date))%>%filter(region!="Airport")

dtafig$region<-factor(dtafig$region, 
                    levels = unique(dtafig$region) 
                    [order(dtafig$cfr, decreasing = TRUE)])

plot_ly(dtafig, x=~region, y=~cfr, type = 'bar')%>% 
    layout(
            title = c("COVID-19 mortality rate by region"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Deaths per 100 confirmed cases (%)"), 
            legend = list(x = 1.2, y = 0.9, 
                          title=list(text='Provinces*'))
            )
```

---

#####__6.2. Trends of case fatality rate by region over time__

```{r,  results='asis', fig.align="left", out.width="800px"}
dtafig<-dtabyregion%>%filter(cfr>0)%>%filter(region!="Airport")

plot_ly(dtafig, x=~date, y=~cfr, 
        type = 'scatter', mode = 'lines', color= ~region, 
        colors = c("dodgerblue3", "blue", "red",  "dodgerblue1")) %>%
        layout(
            title = c("Trend of CFR by region"),
            xaxis = list(title = "Date",  
                         tickfont = list(size=10)),
            yaxis = list(title = "Deaths per 100 cases (%)"),
            legend = list(x = 0.1, y = 0.7, 
                          title=list(text='Provinces*'))
            )
```

---

<p style="color:gray">
See [GitHub](https://github.com/yoonjoung/COVID19_SouthKorea) for data, code, and more information.  
For typos, errors, and questions, contact me at [www.isquared.global](https://www.iSquared.global/YJ). 

_Making Data Delicious, One Byte at a Time_, in good times and bad times.</p>